\documentclass[11pt,letterpaper]{report}
\usepackage{amssymb,amsfonts,color,graphicx,amsmath,enumerate}
\usepackage{tikz}
\usepackage{amsthm}
\usepackage{bbm}
\usepackage{hyperref}

\newcommand{\naturals}{\mathbb{N}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\complex}{\mathbb{C}}
\newcommand{\reals}{\mathbb{R}}
\newcommand{\mcal}[1]{\mathcal{#1}}
\newcommand{\rationals}{\mathbb{Q}}
\newcommand{\Lp}[2]{\left\|{#1}\right\|_{L^{#2}}}
\newcommand{\field}{\mathbb{F}}
\newcommand{\affine}{\mathbb{A}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Prob}{\mathbb{P}}
\newcommand{\Var}{\text{Var}}
\newcommand{\ind}{\mathbbm{1}}
\newcommand{\Cov}{\text{Cov}}

\newenvironment{solution}
{\begin{proof}[Solution]}
{\end{proof}}

\voffset=-3cm
\hoffset=-2.25cm
\textheight=24cm
\textwidth=17.25cm
\addtolength{\jot}{8pt}
\linespread{1.3}

\begin{document}
\noindent{\em Liam Hardiman\hfill{March 2, 2020} }
\begin{center}
{\bf \Large 271B - Homework 4}
\vspace{0.2cm}
\hrule
\end{center}

\noindent\textbf{Problem 1. }
Consider the process
\[
X_t = B^{(1)}_tB^{(2)}_t + e^{t}-t_0 + (B^{(3)}_t)^2,
\]
where $\E[B^{(i)}_tB^{(j)}_t] = \rho_{i,j}t$. Find the stochastic differential equation satisfied by this process.
\begin{solution}
	Let $g(t, x_1, x_2, x_3) = x_1x_2+e^t-t_0+x_3^2$. We apply It\^o's lemma to obtain
	\begin{align*}
		dX_t &= e^tdt + B^{(2)}_tdB^{(1)}_t + B^{(1)}_tdB^{(2)}_t + 2B^{(3)}_tdB^{(3)}_t + (dB^{(3)}_t)^2 + \rho_{1,2}dt\\
		&= (e^t + 1+\rho_{1,2})dt + B^{(2)}_tdB^{(1)}_t + B^{(1)}_tdB^{(2)}_t + 2B^{(3)}_tdB^{(3)}_t
	\end{align*}
\end{solution}


\noindent\textbf{Problem 2. }
Consider
\[
Z_t = \mu_tdt + \theta_t\cdot dB_t,\ Z_0 = z_0,
\]
for $B$ a standard $n$-dimensional Brownian motion and $\theta$ a bounded deterministic $n$-dimensional vector process and $\mu$ a bounded deterministic process, all progressively measurable.
\begin{enumerate}[(a)]
	\item Use the Kolmogorov forward equation to derive the density of $Z_T$.
	\begin{solution}
		Let $p_t(x)$ be the density of $Z_t$. The Kolmogorov forward equation says that
		\[
		\begin{cases}\partial_tp_t(x) = \left(-\mu_t\partial_x + \frac{1}{2}|\theta_t|^2\partial^2_x\right)p_t(x)\\
		p_0(x) = \delta(x - z_0)
		\end{cases}.
		\]
		If $\widehat{p}_t(\omega)$ is the Fourier transform of $p_t(x)$, then after integrating by parts twice we arrive at
		\[
		\begin{cases}\partial_t\widehat{p}_t(\omega) = \left(i\omega\mu_t -\frac{1}{2}\omega^2|\theta_t|^2\right)\widehat{p}_t(\omega)\\
		\widehat{p}_0(\omega) = e^{iz_0\omega}
		\end{cases}.
		\]
		The first equation is a separable ODE, so we obtain
		\[
		\widehat{p}_t(\omega) = e^{iz_0\omega}\exp\left[\int_0^t\left(i\omega\mu_s - \frac{1}{2}\omega^2|\theta_s|^2 \right)ds \right].
		\]
		Taking the inverse Fourier transform gives $p_t(x)$.
	\end{solution}
\end{enumerate}

\noindent\textbf{Problem 3. }
Show that
\[
\int_0^tB_s\circ dB_s = \frac{1}{2}B_t^2.
\]
\begin{proof}
	For all $i$ let $t_i^* = \frac{1}{2}(t_i +t_{i-1})$. We have
	\begin{align*}
		\sum_{i = i}^{t/\Delta t}B_{t_i^*}(B_{t_i}-B_{t_{i-1}}) &= \sum_{i = i}^{t/\Delta t}(B_{t_{i-1}} + (B_{t_i^*}-B_{t_{i-1}}))(B_{t_i}-B_{t_{i-1}})\\
		&= \sum_{i=1}^{t/\Delta t}B_{t_{i-1}}(B_{t_i}-B_{t_{i-1}}) + \sum_{i=1}^{t/\Delta t}(B_{t_i^*} - B_{t_{i-1}})[(B_{t_i^*} - B_{t_{i-1}})+(B_{t_i} - B_{t_i^*})].
	\end{align*}
	The first sum converges to the It\^o integral $\int_0^tB_sdB_s = \frac{1}{2}B_t^2 - \frac{1}{2}t^2$, while the second sum converges to $\frac{1}{2}t^2$. In all, we have that $\int_0^tB_s\circ dB_s = \frac{1}{2}B_t^2$.
\end{proof}

\noindent\textbf{Problem 4. }
Show that the following quantities define metrics on the appropriate spaces.
\begin{gather*}
	[f] = \sum_{n=1}^\infty 2^{-n}\left( 1 \land \sqrt{\E\int_0^nf_s^2\ ds} \right)\\
	\|X\| = \sum_{n=1}^\infty 2^{-n}\left(1 \land \sqrt{\E X_n^2} \right).
\end{gather*}
\begin{proof}
	Symmetry and ``zero if and only if zero'' are obvious, so we check the triangle inequality. This essentially follows from the triangle inequality of the $L^2(\Prob \otimes \text{Leb})$ norm: we recognize $(\E\int_0^n f^2_s\ ds)^{1/2}$ as the $L^2$ norm of $f\cdot \ind_{\Omega \times [0, n]}$, which we write as $\|f\|_{2(n)}$. We then have
	\[
	1 \land \|f+g\|_{2(n)} \leq 1 \land (\|f\|_{2(n)}+\|g\|_{2(n)}).
	\]
	The desired conclusion follows from considering the two cases of $\|f\|_{2(n)}+\|g\|_{2(n)} \leq 1$ and $\|f\|_{2(n)}+\|g\|_{2(n)} \geq 1$. The exact same argument works for $\|X\|$.
\end{proof}

\noindent\textbf{Problem 5. }
Consider
\[
Z_t = \exp\left(\int_0^t\theta_s\cdot dB_s - \frac{1}{2}\int_0^t|\theta_s|^2ds\right),
\]
for $B$ an $n$-dimensional standard Brownian motion and $\theta$ a bounded $n$-dimensional progressively measurable vector process.
\begin{enumerate}[(a)]
	\item Show that $Z$ is a martingale.
	\begin{proof}
		Let $Y_t$ be the expression inside the exponential defining $Z_t$:
		\[
		Y_t = \int_0^t\theta_s\cdot dB_s - \frac{1}{2}\int_0^t|\theta_s|^2ds.
		\]
		This gives
		\[
		dY_t = -\frac{1}{2}|\theta_t|^2dt + \theta_t\cdot dB_t,\quad (dY_t)^2 = |\theta_t|^2.
		\]
		By It\^o's lemma we have
		\begin{align*}
			dZ_t &= Z_tdY_t + \frac{1}{2}Z_t(dY_t)^2\\
			&= Z_t\theta_t\cdot dB_t.
		\end{align*}
		The restrictions on $\theta$ guarantee that $Z_t\theta_t$ is of class II$^*$, so $Z_t$ is an It\^o process with only a martingale component, and is hence a martingale.
	\end{proof}

	\item Assume that $\theta$ is independent of $B$. Derive an expression for the variance of $Z_t$.
	\begin{solution}
	The variance of $Z_t$ is given by
	\[
	\Var[Z_t] = \E[Z_t^2] - \E[Z_t]^2.
	\]
	Since $Z_t$ is a martingale, $\E[Z_t] = \E[Z_0] = 1$. By the It\^o isometry, martingale property, and independence, we have
	\begin{align*}
		\E[Z_t^2] &= \E\left( \int_0^t Z_s\theta_s\cdot dB_s+1\right)^2 \\
		&= \E\left(\int_0^tZ_s\theta_s\cdot dB_s\right)^2 + 2\E\int_0^tZ_s\theta_s\cdot dB_s + 1\\
		&= \E\left[\int_0^tZ_s^2|\theta_s|^2ds\right]+2\sum_{i<j}\E\left(\int_0^tZ_s\theta^{(i)}_sdB^{(i)}_s\right)\E\left(\int_0^tZ_s\theta^{(j)}_sdB^{(j)}_s\right)+1\\
		&= \E\left[\int_0^tZ_s^2|\theta_s|^2ds\right]+1.
	\end{align*}
	Thus, we have
	\[
	\Var[Z_t] = \E\left[\int_0^tZ_s^2|\theta_s|^2ds\right].
	\]
	\end{solution}
\end{enumerate}

\noindent\textbf{Problem 6. }
Consider the scalar process
\[
dZ_t = \mu_tdt + dB_t,
\]
for $\mu_t$ bounded and progressively measurable. Find $X_t$ so that $Z_tX_t$ is a martingale.
\begin{solution}
	Let $X_t$ be defined by
	\[
	X_t = \exp\left(-\int_0^t\mu_t\ dB_t - \frac{1}{2}\int_0^t\mu_t^2\ dt \right).
	\]
	By It\^o's lemma and the same reasoning used in the previous problem, we have
	\[
	dX_t = -X_t\mu_t\ dB_t.
	\]
	We also have
	\begin{align*}
		d(Z_tX_t) &= Z_tdX_t + X_tdZ_t + d\langle Z_t, X_t\rangle\\
		&= Z_t(-X_t\mu_tdB_t) + X_t(\mu_tdt + dB_t) - X_t\mu_tdt\\
		&= X_t(1-Z_t\mu_t)dB_t.
	\end{align*}
	This is an It\^o process with only a martingale component, so $Z_tX_t$ is a martingale.
\end{solution}

\end{document}