\documentclass[11pt,letterpaper]{report}
\usepackage{amssymb,amsfonts,color,graphicx,amsmath,enumerate}
\usepackage{tikz}
\usepackage{amsthm}
\usepackage{bbm}
\usepackage{hyperref}

\newcommand{\naturals}{\mathbb{N}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\complex}{\mathbb{C}}
\newcommand{\reals}{\mathbb{R}}
\newcommand{\mcal}[1]{\mathcal{#1}}
\newcommand{\rationals}{\mathbb{Q}}
\newcommand{\Lp}[2]{\left\|{#1}\right\|_{L^{#2}}}
\newcommand{\field}{\mathbb{F}}
\newcommand{\affine}{\mathbb{A}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Prob}{\mathbb{P}}
\newcommand{\Var}{\text{Var}}
\newcommand{\ind}{\mathbbm{1}}
\newcommand{\Cov}{\text{Cov}}

\newenvironment{solution}
{\begin{proof}[Solution]}
{\end{proof}}

\voffset=-3cm
\hoffset=-2.25cm
\textheight=24cm
\textwidth=17.25cm
\addtolength{\jot}{8pt}
\linespread{1.3}

\begin{document}
%\noindent{\em Liam Hardiman\hfill{Date} }
\begin{center}
{\bf \Large Math 175 - Homework 3}
\vspace{0.2cm}
\hrule
\end{center}

\begin{enumerate}
	\item There are twenty-five beads in a bag, each colored red, blue or green. Prove that there are at least nine beads of the same color.
	\vfill

	\item Let $A = \{a_1, a_2, a_3, a_4, a_5\}$ be a set of five positive integers. Show that for any permutation $a_{i_1}, a_{i_2}, a_{i_3}, a_{i_4}, a_{i_5}$ of $A$, the product
	\[
	(a_{i_1}-a_1)(a_{i_2}-a_2)(a_{i_3}-a_3)(a_{i_4}-a_4)(a_{i_5}-a_5).
	\]
	is always even.
	\vfill
	\item Let $K_n$ be the complete graph on $n$ vertices, i.e. the graph with $n$ vertices where every pair of vertices is connected by an edge. 
	\begin{enumerate}
		\item Suppose we color each edge of $K_6$ either red or blue. Prove that there is at least one monochromatic triangle. A triangle is monochromatic if all of its edges have the same color.
		\item Is part (a) still true for $K_5$?
		\item Prove that in any coloring of $K_6$ we actually have at least \textit{two} monochromatic triangles. \textit{Hint: Count the complement.}
	\end{enumerate}
	\vfill

	\item \begin{enumerate}
		\item Is it possible to cover a chessboard (an 8 by 8 grid where squares alternate between two colors) with dominoes that cover exactly two squares?
		\item What if we delete two white squares?
	\end{enumerate}

	\vfill
	
	\item \begin{enumerate}
		\item Let $A$ be a set of $m$ positive integers where $m\geq 1$. Show that there exists a nonempty subset $B$ of $A$ such that the sum $\sum_{x\in B}x$ is divisible by $m$.
		\item Let $X\subseteq \{1, 2, \ldots, 99\}$ and $|X| = 10$. Show that it is possible to select two disjoint nonempty proper subsets $Y, Z$ of $X$ such that $\sum_{y\in Y}y = \sum_{z\in Z}z$.
	\end{enumerate}
	\vfill
\end{enumerate}

\end{document}