\documentclass[11pt,letterpaper]{report}
\usepackage{amssymb,amsfonts,color,graphicx,amsmath,enumerate}
\usepackage{amsthm}

\newcommand{\naturals}{\mathbb{N}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\complex}{\mathbb{C}}
\newcommand{\reals}{\mathbb{R}}
\newcommand{\mcal}[1]{\mathcal{#1}}
\newcommand{\rationals}{\mathbb{Q}}
\newcommand{\field}{\mathbb{F}}
\newcommand{\Var}{\text{Var}}
\newcommand{\ind}{\mathbbm{1}}
\newcommand{\Cov}{\text{Cov}}

\newenvironment{solution}
{\begin{proof}[Solution]}
{\end{proof}}

\voffset=-3cm
\hoffset=-2.25cm
\textheight=24cm
\textwidth=17.25cm
\addtolength{\jot}{8pt}
\linespread{1.3}

\begin{document}
\begin{center}
{\bf \Large Math 130B - Homework 3}
\vspace{0.2cm}
\hrule
\end{center}


\begin{enumerate}
	\item Let $Z$ be a standard normal random variable and let $g$ be a differentiable function with derivative $g'$.
	\begin{enumerate}
		\item Show that $E[g'(Z)] = E[Zg(Z)]$. What assumptions do you need in order for this to be true?
		\item Show that $E[Z^{n+1}] = nE[Z^{n-1}]$.
		\item Find $E[Z^4]$.
		\item Find $E[Z^{2n}]$ and $E[Z^{2n+1}]$.
		\item Let $X\sim \mcal{N}(\mu, \sigma^2)$. Show that $E[(X-\mu)g(X)] = \sigma^2E[g'(X)]$ when both sides exist.
	\end{enumerate}

	\item Let $\phi(x)$ be the standard normal density and let $\Phi(x)$ be the standard normal cdf. Show that $\phi'(x) + x\phi(x) = 0$. Deduce that
	\[
	\frac{1}{x} - \frac{1}{x^3} < \frac{1-\Phi(x)}{\phi(x)} < \frac{1}{x} - \frac{1}{x^3} + \frac{3}{x^5},\quad x>0.
	\]
	This inequality is useful since it helps estimate $\Phi(x)$, which doesn't have a closed-form expression.

	\item Four witnesses, Alice, Bob, Carol, and Dave, at a trial each tell the truth with probability $\frac{1}{3}$ independent of each other. In their testimonies, Alice claimed that Bob denied that Carol declared that Dave lied. What is the conditional probability that Dave told the truth?

	\item Consider a stick of unit length. Break the stick at two points chosen uniformly at random. What is the probability that you can form a triangle from the three pieces that remain.

	\item A roulette wheel has 38 slots: 18 red, 18 black, and 2 green. A ball is tossed into the wheel and eventually settles in a slot at random. You play many games, betting \$1 on red each time. If the ball lands in a red slot you win \$1, otherwise you lose the dollar you bet. After $n$ games, what is the probability that you have more money than you started with? Give (approximate) numerical answers for $n=100$ and $n=1000$. How would the answer change if there were no green slots?

	\item How many times do you need to toss a fair coin in order to get 100 heads with probability at least 0.9?

\end{enumerate}



\end{document}