
\documentclass[12pt]{article}  
\usepackage{amsmath}
\usepackage{amssymb}

\pagestyle{empty}

\unitlength 1cm
\textheight 22cm
\textwidth 17cm
\oddsidemargin -0.5cm
\evensidemargin -0.5cm
\topmargin -1.5cm
\topskip 0cm
\headheight 0.5cm
\headsep 1cm
\marginparwidth 1.2cm

\begin{document}

\begin{center}
	\textbf{Math 2B Worksheet: 5.1 Areas and Distances}
\end{center}

\emph{Write your names and Student ID numbers at the top of the page}


\begin{enumerate}
\item Consider the curve $y=x^2$ on the interval $[1,3]$
\begin{enumerate}
	\item Is $y$ increasing or decreasing?\\
	
	\vspace{1cm}
	
	\item Estimate the area under $y$ on $[1,3]$ using left endpoints and 4 rectangles.  Is your estimate an under-estimate or an over-estimate?\\
	
	\vfill
	
	\item Repeat part (b) using right endpoints.  Is this an under-estimate or an over-estimate?\\
	
	\vfill
	
	\item If a function $f$ is increasing on $[a,b]$, then based upon your answers above, choose "Over" or "Under" in each part below.
	\begin{enumerate}
		\item Estimating the area under $f$ using left endpoints will result in an (Over/Under) estimate.
		\item Estimating the area under $f$ using right endpoints will result in an (Over/Under) estimate
	\end{enumerate}
	
	
	\newpage
	
	\item  What do you think will happen when estimating the area under the curve of a function $g$ that is \textbf{\textit{decreasing}} on $[a,b]$? Test this by estimating the area under $g(x)=6-x^2$ on $[0,2]$ using both left and right endpoints and 4 rectangles.
\end{enumerate}

\vfill

\item The speed of a runner increased steadily during the first
three seconds of a race. Her speed at half-second intervals
is given in the table. Find lower and upper estimates for
the distance that she traveled during these three seconds.
\\
\vspace{.5cm}
\begin{tabular}{| p{2cm} | p{.75cm}| p{.75cm}|p{.75cm}|p{.75cm}|p{.75cm}|p{.75cm}|p{.75cm}|}
	\hline
	$t\; (s)$ & $0.0$ & $0.5$ & $1.0$ & $1.5$ & $2.0$ & $2.5$ & $3.0$ \\
	\hline
	$v\;\text{(ft/s)}$ & $0$ & $6.2$ & $10.8$ & $14.9$ & $18.1$ & $19.4$ & $20.2$ \\
	\hline
	\end {tabular}
	
	\vfill
	
	
% 	\vspace{7cm}
% 	\item Oil leaked from a tank at a rate of r(t) liters per hour. The rate
% 	decreased as time passed and values of the rate at two-hour
% 	time intervals are shown in the table. Find lower and upper
% 	estimates for the total amount of oil that leaked out.
% 	\\
% 	\vspace{.5cm}
% 	\begin{tabular}{| p{1.75cm} | p{.75cm}| p{.75cm}|p{.75cm}|p{.75cm}|p{.75cm}|p{.75cm}|}
% 		\hline
% 		$t\;\text{(h)}$ & $0$ & $2$ & $4$ & $6$ & $8$ & $10$\\
% 		\hline
% 		$r(t)\;\text{(L/h)}$ & $8.7$ & $7.6$ & $6.8$ & $6.2$ & $5.7$ & $5.3$\\
% 		\hline
% 	\end {tabular}
	


\end{enumerate}



\end{document}

