\documentclass[11pt,letterpaper]{report}
\usepackage{amssymb,amsfonts,color,graphicx,amsmath,enumerate}
\usepackage{tikz}
\usepackage{amsthm}

\newcommand{\naturals}{\mathbb{N}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\complex}{\mathbb{C}}
\newcommand{\reals}{\mathbb{R}}
\newcommand{\mcal}[1]{\mathcal{#1}}
\newcommand{\rationals}{\mathbb{Q}}
\newcommand{\Lp}[2]{\left\|{#1}\right\|_{L^{#2}}}
\newcommand{\field}{\mathbb{F}}
\newcommand{\affine}{\mathbb{A}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Prob}{\mathbb{P}}
\newcommand{\Var}{\text{Var}}
\newcommand{\ind}{\mathbbm{1}}
\newcommand{\Cov}{\text{Cov}}

\newenvironment{solution}
{\begin{proof}[Solution]}
{\end{proof}}

\voffset=-3cm
\hoffset=-2.25cm
\textheight=24cm
\textwidth=17.25cm
\addtolength{\jot}{8pt}
\linespread{1.3}

\begin{document}
\begin{center}
{\bf \Large Math 175 - Graph Theory Exercises}
\vspace{0.2cm}
\hrule
\end{center}

\begin{enumerate}
	\item \begin{enumerate}[(a)]
		\item Show that every tree with maximum degree $k$ has at least $k$ leaves.
		\item Which such trees have exactly $k$ leaves?
	\end{enumerate}

	\vfill

	\item A saturated hydrocarbon is a molecule $C_mH_n$ in which every carbon atom $C$ has four bonds, every hydrogen atom $H$ has one bond, and no sequence of bonds forms a cycle. Show that if $C_mH_n$ is a saturated hydrocarbon, then $n = 2m+2$.
	\vfill

	\item A \textbf{center} of a graph $G = (V, E)$ is a vertex $u$ such that $\max\{d(u, v): v\in V\}$ is as small as possible, where $d(u, v)$ is the length of a shortest path connecting $u$ and $v$ if such a path exists and $\infty$ otherwise.
	\begin{enumerate}
		\item Let $T$ be a tree on at least three vertices and let $T'$ be the tree obtained from $T$ by deleting all of its leaves. Show that $T$ and $T'$ have the same centers.
		\item Deduce that every tree has either exactly one center or two adjacent centers.
	\end{enumerate}

	\vfill

	\item Show that at any party with $n\geq 2$ people, there are two people who know the same number of people.

	\vfill

	\item Show that in any graph $G$, the distance function is a \textbf{metric}. That is, show that
	\begin{enumerate}
		\item $d(u,v) = 0$ if and only if $u=v$.
		\item $d(u,v) = d(v,u)$.
		\item $d(u,v)+d(v,w)\geq d(u, w)$.
	\end{enumerate}

	\vfill

	\item An edge $e$ of a graph $G$ is a \textbf{cut edge} if its deletion causes the number of connected components to increase. That is, $e$ is a cut edge of $G$ if $G\setminus e$ has one more connected component than $G$. Show that $e$ is a cut edge of $G$ if and only if $e$ belongs to no cycle of $G$.
	\vfill

\end{enumerate}


\end{document}